---
title: Enare
description: |
  Greetings from Allie
author:
- name: Allie Lyamuya
date: 07-12-2021
categories:
- homework 6
- enare
- allie lyamuya
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction to Class
Hello everyone! My name is Alphoncina (Allie) and I'm a master of public policy
and administration student here at UMass. I'm interested in ways that data and 
data science can be used to facilitate policy making and informing policymakers. 


## Statistical Analysis in R

```{r echo=TRUE, results='hide'}
library(tidyverse)
library(nycflights13)
library(ggplot2)
```

```{r}
# Load some data from the `nycflights13` package
flights <- flights
airlines <- airlines

# Inspect the data
colnames(flights)
colnames(airlines)

str(flights)
str(airlines)

dim(flights)
dim(airlines)
```

```{r}
# Creating a histogram for `flights$arr_delay`
hist(flights$arr_delay, 
     main = "Histogram of Arrival Delay",
     xlab = "flights$arr_delay",
     col = "light green")

# select variables from the flights data
flightsSelected <- flights %>%
  dplyr::select(carrier, dep_time, dep_delay, arr_time, arr_delay, flight, 
                origin, dest, air_time, distance)
flightsSelected
```


***
Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


