---
title: "efleming_hw1"
description: |
  Short introductory analysis on covid responses
author:
  - name: Emerson Fleming
    url: http://umass.edu/sbs/dacss
date: 07-21-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
```{r setup, include=FALSE}
require(tidyverse)
library(dplyr)
myDataPath <- read.csv("../../_data/covid-responses.tab - covid-responses.tab.csv", header=TRUE) 
covid_responses <- myDataPath
```
**COVID Responses Table**
```{r}
view(myDataPath) %>%
  slice(1:10) 
```
This is the parent data set. Overall, we can see it is quite messy. I am sure the gather and spread functions could be used to tidy the data but this will not be necessary for the portion of data I will examine.

1. The following data has 3460 rows and 80 different columns.
2. Variables include "reflection", "progress", "handshake", and many others.

```{r}
myDataPath <- read.csv("../../_data/covid-responses.tab - covid-responses.tab.csv", header=TRUE)
covid_responses <- myDataPath
covid_responses %>% #addressed my data set directly to begin consolidation efforts
  select(social, handshake, stores, curfew, perceivedreaction) %>%  #implemented the following variables into my set
    na.omit() %>% #omitted all of the "NA" variables.
    slice(1:20) #used the following command to print the first 25 rows of my table
```
I did feel the data could be consolidated better to begin building an argument. For this example, I attempted to search for a correlation between "perceived reaction" to the Corona virus outbreak and whether or not participants engaged in handshakes, participated in social endeavors, traveled to stores, and abode by curfew hours. Overall, I was unable to deduce anything worthwhile. However, I was able to practice my tidying skills. Until next time.

