---
title: Homework One
description: |
  This assignment is writing R script to get data into R. The dataset used for this assignment is scdb.
author:
- name: Kristina Becvar
date: 07-14-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(blogbuilder)
library(distill)
library(tidyverse)
iamstudent()
```

## **Reading in the data:**

```{r echo=TRUE}
scdb.data <- read.delim("scdb.csv", sep = ",", header  = TRUE)
```

## **About the data command:**

+ This object is named "scdb.data" because it holds data from the Supreme Court Database (scdb)
+ The data file for scdb in my working directory
+ The delimiter is "," because the file is a CSV (comma-separated value)
+ The header is "TRUE" because there is a header in the data with variable names

## **Checking the first few rows**

+ The head() function returns the variable names & the first 6 rows of the data file by default
+ head(scdb.data) previewed so many columns that the knitted post displayed them poorly. So I did some searching after completing the next steps to adjust how it displayed on my post.
+ I looked up how to select just one column to preview, "caseName",and executed it below.

```{r echo=TRUE}
head(scdb.data$caseName)
```
## Checking the number of rows of data and the number of columns in the dataset

```{r echo=TRUE}
dim(scdb.data)
```
+ There are 8966 rows in this database
+ There are 53 columns in the database

## Checking the column names

```{r echo=TRUE}
colnames(scdb.data)
```

## Determining the use of the gather() and spread() functions

+ To determine what to do with the data next, I used the "as_tibble" function to see the data in that format.

```{r echo=TRUE}
(scdb.tidy <- as_tibble(scdb.data))
```

+ At this point, I'm unsure how I would use the functions with this dataset

## Executing an operation that requires creating a new object and doing something to that object
+ I selected the categories "term" and "decisionDirection" 

```{r echo=TRUE}
scdb.tidy2 <- select(scdb.tidy,term, decisionDirection)
```

+ Then, I filtered the term date to only include decisions from term year 1976

```{r echo=TRUE}
(scdb.1976 <- filter(scdb.tidy2, term == "1976"))
```
