---
title: Homework One
description: |
  This assignment is writing R script to get data into R. The dataset used for this assignment is scdb.
author:
- name: Kristina Becvar
date: 07-14-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## **Reading in the data:**

```{r echo=TRUE}
scdb.data <- read.delim("scdb.csv", sep = ",", header  = TRUE)
```

## **About the data command:**

+ This object is named "scdb.data" because it holds data from the Supreme Court Database (scdb)
+ The data file for scdb in my working directory
+ The delimiter is "," because the file is a CSV (comma-separated value)
+ The header is "TRUE" because there is a header in the data with variable names

## **Checking the first few rows**

+ The head() function returns the variable names & the first few rows of the data file
+ In this case, the preview of 6 lines of 53 columns were quite large, so I looked up how to select just one column to preview, "caseName".

```{r echo=TRUE}
head(scdb.data$caseName)
```

## Checking the number of rows of data and the number of columns in the dataset

```{r echo=TRUE}
dim(scdb.data)
```

## The dimensions of the database are 8966 rows by 53 columns

## Checking the column names

```{r echo=TRUE}
colnames(scdb.data)
```

## Prompt on the gather() and spread() functions

## Executing an operation that requires creating a new object and doing something to that object

## Executing a creative and fun manipulation to a portion of these data that hasn't been covered so far in the course materials



