---
title: "efleminghw3"
author: "Emerson Fleming"
date: "7/23/2021"
output: html_document
---
```{r setup, include=FALSE}
require(tidyverse)
require(tidyr)
require(dplyr)
```

```{r}
head(myDataPath[,1:16], 10)
```
The first problem with the dataset that needed to be addressed was for the missing values in the dataset.
So far, there are several key issues that have been addressed and will be included at a later juncture:
1. I am not sure which years were used to compile this data
2. I am not sure whether this data just includes sales in the US or worldwide sales
3. I am not sure whether this data was compiled using multiple years or one single year (as my research on the data will indicate)

As we can see, there appears to be a missing value for the resale value of the BMW 323i. The last year the BMW 323i was in year 2000 (which is about the year that I assume this data is based around according to my research). The resale price for the BMW 328i is listed at $28,675. O

source
(https://www.cars.com/research/bmw-323/)
(https://www.cars.com/research/bmw-323/)
