---
title: "Genovese HW1 Revision 7.27.21"
description: |
  Here is my revision for HW1
author:
  - name: Curtis Atkisson
    url: http://umass.edu/sbs/dacss
date: 07-27-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
```{r, include=FALSE}
library(tidyverse)
```
**Loading in the Data**
```{r}
myData <- read.delim("../../_data/congressional_hearings.csv", sep = ",", header=T)
```
I loaded my data into the variable, myData using read.delim and finding the 
correct source file.  
**Finding basic information on data**
```{r}
head(myData)
nrow(myData)
ncol(myData)
colnames(myData)
```
I found the basic information on the data I just read in. Using head, I found 
the first few values for each column, while also finding how many columns and 
rows there were. Finally, I listed the column names.
**Tidying data**
```{r}
tidymyData <- gather(myData, key = "Chamber", value = "ChamberNumber", 6:8)
```
I tidied the data to gather the chamber number columns into a column that 
gives me the data of where each bill originated.
**Finding Median**
```{r}
class(myData$Sessions)
sessionMedian <- median(myData$Sessions, na.rm = TRUE)
print(sessionMedian)
```
I found the median, which tells us the average amount of sessions it took to 
pass a bill in Congress.
**Finding Mode**
```{r}
class(myData$Chamber)
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x,ux)))]
}
Mode(myData$Chamber)
```
Using a mode function I found on stack overflow, I was able to see which 
chamber originated the most amount of bills. Using that function, we can see 
that the House of Representatives originated the most amount of bills in 
Congress.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


