---
title: "HW2"
description: |
  Data exploration and research question formulation for final project
author:
  - name: Allie Lyamuya
date: 07-23-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Overview

For my final project, I have identified several data sets from the [World Bank 
Open Data](https://data.worldbank.org) 
Since there was no single data set that included all the variables I need to 
analyze the research questions that I have, I decided to join several data sets 
together. I obtained six different data sets from the World Bank, then I used 
the data wrangling tools that we have so far learned in class to tidying the 
data. Because all the data sets came from the same source, joining
them was not as challenging. You can learn more about the whole [tidying process
here](https://rpubs.com/alyamuya/793043) 


## Research Question 

Does the country's level of economic development determine the level of 
information communications technologies (ICTs) usage? 

* I will be analyzing this question based on the dataset I recreated which
include the following variables: country name, year, individual internet use, 
population, gdp per capita, fixed broadbands subscription, cellular mobile
subscription, telephone subscription. The dataset include data from 2007 to 2017. 

## Data Exploration 

```{r}
# Importing data
library(readr)
myData <- read.csv("../../_data/digitalConnectivityData.csv")

head(myData)
```

```{r}
library(tidyverse)
library(dplyr)

# Describing data
glimpse(myData)
```


## Variable Identification

```{r}
# Check for missing values
is.na(myData$internet_usage)
is.na(myData$gdp_percapita)
is.na(myData$fixed_broadbands)
is.na(myData$mobile_cellular)
is.na(myData$telephone)

# Position of missing values
which(is.na(myData$internet_usage))
which(is.na(myData$gdp_percapita))
which(is.na(myData$fixed_broadbands))     
which(is.na(myData$mobile_cellular))
which(is.na(myData$telephone)) 
```

<br>

```{r}
# recode 0 to missing values in data set
myDataNew <- myData
myDataNew[is.na(myDataNew)] <- 0

# Double check for missing values in a new data set
which(is.na(myDataNew))

```


```{r}
options(scipen = 999, digits = 4)
# Finding the range of values in relevant variables 
range(myDataNew$internet_usage)
range(myDataNew$gdp_percapita)
range(myDataNew$fixed_broadbands)
range(myDataNew$mobile_cellular)
range(myDataNew$telephone)

# Check for the distribution of values in variables 
summary(myDataNew$internet_usage)
summary(myDataNew$gdp_percapita)
summary(myDataNew$fixed_broadbands)
summary(myDataNew$mobile_cellular)
summary(myDataNew$telephone)

```

<br>

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


