---
title: "SCortez HW 1"
description: |
  Homework 1
author: Stephanie Cortez
date: 07-26-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# I am using a dataset I found in LAUSD website on SENIS (School Equity Needs Index Scores) for elementary schools. File is SENIScoresElementaryShools-20212022.xlsx.
```{r}
#Importing data
library(readxl)
SENIScoresElementaryShools_20212022 <- read_excel("~/UMASS/DACSS/601/GitHub/DACSS601SS22021/_data/SENIScoresElementaryShools-20212022.xlsx")
SENISData <- SENIScoresElementaryShools_20212022
SENISData
```

#Now, let's check the first few rows. I'll let you remember how to do that!
```{r}
head(SENISData)
```
#How many rows of data are there in this dataset? How many columns are there in the dataset?
```{r}
dim(SENISData)
```
#What are the column names?
```{r}
colnames(SENISData)
```

#Should any of these columns have gather() or spread() applied to them to make them more useful? If so, perform those operations. If you want a challenge, try to use
p#ivot_longer() or pivot_wider() instead.
```{r}
pivot_wider(SENISData, "Local District")
```
#Do one operation that requires creating a new object and doing something to that object
```{r}
#Assigning new tidyset as tidySENItest that shows columns Local District, Homeless Count and Foster Count from SENISData
tidySENItest <- select(SENISData, `Local District`, `Homeless Count`, `Foster Count`)
head(tidySENItest)
```
#Do one thing to some portion of these data that hasn't been covered so far in the course materials. Be creative and have fun!

``````{r}
#Pulling [x,y] value
SENISData[5,10]
```

``````{r}
#Pulling/Printing first 10 rows and all columns
print(SENISData, n=10, width = Inf)
```


