---
title: SCortez HW 1
description: |
  Homework 1
author: Stephanie Cortez
date: 07-26-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---
# Homework 1
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE)
```

### I am using a dataset I found in LAUSD website on SENIS (School Equity Needs Index Scores) for elementary schools. File is SENIScoresElementaryShools-20212022.xlsx.  
**Importing data**
```{r, include=TRUE}
#Importing data
library(readxl)
SENIScoresElementaryShools_20212022 <- read_excel("../../_data/SENIScoresElementaryShools-20212022.xlsx")
SENISData <- SENIScoresElementaryShools_20212022
SENISData
```

### Now, let's check the first few rows. I'll let you remember how to do that!
**Ran head() function**
```{r, include=TRUE}
head(SENISData)
```
### How many rows of data are there in this dataset? How many columns are there in the dataset?
**Ran dim() function**
```{r,include=TRUE}
dim(SENISData)
```
### What are the column names?
**Ran colnames() function**
```{r, include=TRUE}
colnames(SENISData)
```

### Should any of these columns have gather() or spread() applied to them to make them more useful? If so, perform those operations. If you want a challenge, try to use pivot_longer() or pivot_wider() instead.
**Ran pivot_wider() function on "Local District" variable to view local district names**
```{r, include=TRUE}
pivot_wider(SENISData, "Local District")
```
### Do one operation that requires creating a new object and doing something to that object
**Assigning new tidyset as tidySENItest that shows columns Local District, Homeless Count and Foster Count from SENISData**
```{r, include=TRUE}
#Assigning new tidyset as tidySENItest that shows columns Local District, Homeless Count and Foster Count from SENISData
tidySENItest <- select(SENISData, `Local District`, `Homeless Count`, `Foster Count`)
head(tidySENItest)
```
### Do one thing to some portion of these data that hasn't been covered so far in the course materials. Be creative and have fun!
**Pulling [x,y] value**
``````{r, include=TRUE}
#Pulling [x,y] value
SENISData[5,10]
```
**Pulling/Printing first 10 rows and all columns using print() function**
``````{r, include=TRUE}
#Pulling/Printing first 10 rows and all columns
print(SENISData, n=10, width = Inf)
```


