---
title: "AG HW 4"
description: |
  This is Alex Genovese's HW 4: Data Visualization
author:
  - name: Alexander Genovese
date: 08-03-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
```{r, include = FALSE}
library(tidyverse)
library(readxl)
```

# Loading in the Data and Creating the Visualization
## Loading in the Excel Sheets
```{r}
UniPerCountry <- read_excel("../../_data/UniPerCountry.xlsx")
GDPPerCountry <- read_excel("../../_data/WorldBankGDPData.xlsx")
```  

## Cleaning the Individual Sheets
```{r}
  names(UniPerCountry)[1] <- 'Country'
  names(UniPerCountry)[2] <- 'Number of Universities'
  GDPPerCountry[1] <- NULL
```  

## Merging the two sheets and cleaning new sheet
```{r}
UnivsGDP <- merge(GDPPerCountry, UniPerCountry,by=c("Country"))
NumberOfUniversities <- UnivsGDP$`Number of Universities`
GDPData <- UnivsGDP$`Millions of US Dollars`/1000000
NumberOfUniversities<- as.numeric(NumberOfUniversities)
```  
For some reason, the vector **NumberofUniversities** was a character class, so I had to change it to numeric using the function, **as.numeric**. 

## Generating Visualization
```{r}
ggplot(UnivsGDP, aes(x = NumberOfUniversities, y = GDPData)) + 
  geom_point() + 
  labs(x = "Number of Universities", y = "GDP (in Millions of US Dollars)", title = "Number of Universities vs. GDP (in Millions of US Dollars)") + 
  geom_label(aes(label=Country),hjust=0, vjust=0) +
  geom_smooth(method = "lm", se = FALSE, color='red')
```  

# Question 1: What the Visualization Demonstrates
I wanted to look at how the number of Universities in a country correlates with the country's GDP. It may seem common sense, but I wanted to have concrete data to see a relationship between education (specifically secondary education) and the country's economy measured in GDP. As such, the independent variable, or the x-axis, is the vector **NumberofUniversities** and the dependent variable, or the y-axis, is the vector **GDPData**. 

# Question 2: Why you chose the visualization approach you selected, and what alternative approaches you considered but decided not to pursue

I chose to use a scatterplot or **geom_point()** because it allows me to look at the large range of data that I am analyzing. It gives both focus to the closer ranged variables as well as the outliers. I could have used other plots such as a quantile plot or a jitterplot, but I believe that the scatterplot would be more appropriate as the outliers would significantly impact the quantile plot, rendering inaccurate data and there are only 25 observations, making a jitterplot essentially the same as a scatterplot.

# Question 3: What you wished, if anything, you could have executed but found limited capability to do.

I would want to adjust the scaling of the visualization to make sure that all the labels could be read, or adjust the positioning of the labels. I would also be interested in adding the linear regression information in the form of a legend. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


