---
title: "Homework 2"
description: |
  For the 2nd homework assignment, I reuse the Congressional dataset and plug my work into a markdown file.
author:
  - name: Alexander Hong
date: 03-31-2021
categories:
  - homework 2
  - congress
  - alex hong
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

For the 2nd homework assignment, I reuse the Congressional dataset and plug my work into a markdown file.

```{r ch, echo=TRUE, results='hide', warning=FALSE, message=F}
library(readr)
ch <- read_csv("../DACSS601 Data/congressional_hearings.csv")
```

## Data Display and Writeup

```{r ch_data, echo=TRUE, warning=FALSE, message=F}
#columns and rows for the dataset
c_col <- ncol(ch)
c_row <- nrow(ch)
year_obs <- tail(table(ch$year))
a <- knitr::kable(year_obs, col.names = c("Year", "Frequency"))

```

The Congressional Hearings dataset contains in existence, all House, Senate, and Joint committees, subcommittees, and select and special committees from 1947 to 2010.
For this dataset, the number of variables in the Congressional dataset is `r c_col` and the number of observations in the dataset is `r c_row`. 


The number of Congressional Hearings for the last five years can be seen below

`r a`