---
title: Homework One
description: |
  This assignment is writing R script to get data into R. The dataset used for this assignment is scdb.
author:
- name: Kristina Becvar
date: 07-14-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(blogbuilder)
library(distill)
library(tidyverse)
iamstudent()
```

## **Reading in the data:**

```{r echo=TRUE}
scdb.data <- read.delim("scdb.csv", sep = ",", header  = TRUE)
```

## **About the data command:**

+ This object is named "scdb.data" because it holds data from the Supreme Court Database (scdb)
+ The data file for scdb in my working directory
+ The delimiter is "," because the file is a CSV (comma-separated value)
+ The header is "TRUE" because there is a header in the data with variable names

## **Checking the first few rows**

+ The head() function returns the variable names & the first 6 rows of the data file by default
+ head(scdb.data) previewed so many columns that the knitted post displayed them poorly. So I did some searching after completing the next steps to adjust how it displayed on my post.
+ I looked up how to select just one column to preview, "caseName",and executed it below.

```{r echo=TRUE}
head(scdb.data$caseName)
```
## **Checking the number of rows of data and the number of columns in the dataset**

```{r echo=TRUE}
dim(scdb.data)
```
+ There are 8966 rows in this database
+ There are 53 columns in the database

## **Checking the column names**

```{r echo=TRUE}
colnames(scdb.data)
```

## **Determining the use of the gather() and spread() functions**

+ To determine what to do with the data next, I used the "as_tibble" function to see the data in that format.

```{r echo=TRUE}
(scdb.tidy <- as_tibble(scdb.data))
```

+ At this point, I don't see a purpose for using those functions with this dataset. I'm almost positive I'm missing an obvious answer but I haven't quite seen it.

## **Executing an operation that requires creating a new object and doing something to that object**

+ I selected the categories "term" and "decisionDirection" 
+ Then, I filtered the term date to only include decisions from term year 1976, randomly chosen because it's the year I was born.
+ Next, I used the 'mutate' function to create a new column using the integers from 'decisionDirection'. 
+ The new column is titled 'ideology', representing either conservative, liberal decisions
+ Finally, I renamed the integers in the ideology column to represent their coded meanings (1 = conservative, 2 = liberal, 3 = unspecified)

```{r echo=TRUE}
scdb.tidy2 <- select(scdb.tidy,term, decisionDirection)
scdb.1976 <- filter(scdb.tidy2, term == "1976") 
(scdb.1976 <- scdb.1976 %>% 
  mutate(ideology = case_when(
    decisionDirection == 1 ~ "conservative",
    decisionDirection == 2 ~ "liberal",
    decisionDirection == 3 ~ "unspecified"
  )))
```

## **Doing one thing to some portion of these data that hasn't been covered so far in the course materials**

+ I wanted to look at graphics of the ideology trend of cases in the supreme court term of 1976. I used ggplot to create a bar chart. 

```{r echo=TRUE}
ggplot(scdb.1976, aes(x = ideology)) + 
  geom_bar(fill = "slategray3", 
           color="black") +
  labs(x = "Ideology", 
       y = "Frequency", 
       title = "Ideology of Supreme Court Decisions in 1976")
```

+ That got me wanting to try something different, and look up how to display it in a different color and with percentages instead of volume on the y axis.

```{r echo=TRUE}
library(dplyr)
library(scales)
plotdata <- scdb.1976 %>%
  count(ideology) %>%
  mutate(pct = n / sum(n),
         pctlabel = paste0(round(pct*100), "%"))
```

```{r echo=TRUE}
ggplot(plotdata, 
       aes(x = reorder(ideology, -pct),
           y = pct)) + 
  geom_bar(stat = "identity", 
           fill = "darkred", 
           color = "black") +
  geom_text(aes(label = pctlabel), 
            vjust = -0.25) +
  scale_y_continuous(labels = percent) +
  labs(x = "Ideology", 
       y = "Frequency", 
       title = "Ideology of Supreme Court Decisions in 1976")
```

+ I can definitely see how the process of visualizing data can be fun once I understand the language better.
