---
title: "Lyamuya_HW1"
description: |
  This is my homework #1 for session 4 
author:
  - name: Allie Lyamuya
date: 07-14-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

For this assignment, I began by creating an R Script in which I imported the
data into R and went on to respond to the included prompts. Below is the list of
codes my R script comprised of:

```{r}
# Run the library
library(tidyverse)

# Import the data into R
hearingsData <- read.csv("/Users/allielyamuya/Python, Data Viz, R/DACSS601SS22021/congressional_hearings.csv")

```

<br>

```{r}
# To get the first few rows of data
head(hearingsData)

# To gain some familiarity with the data
colnames(hearingsData)
dim(hearingsData)
```

<br>

Should any of these columns have gather() or spread() applied to them to make 
them more useful? If so, perform those operations. If you want a challenge, 
try to use pivot_longer() or pivot_wider() instead.

* Given that each "id" and "source" are quite distinct, I don't think `gather()`
and `spread()` or `pivot_longer()` or `pivot_wider()` could have been useful in
this data. 

<br>


```{r echo=T, results='hold'}
# creating a new object and make it to function 
selectedHearingsData <- hearingsData %>%
  dplyr::rename("cisyear" = "CISYear", "filterHouse" = "filter_House", 
                "filterSenate" = "filter_Senate", "filterJoint" = "filter_Joint",
                "cisCommittee" = "CISCommittee") %>%
  dplyr::select(id, source, cisyear, filterHouse, filterSenate, filterJoint,
                cisCommittee)

# Arrange the selectedHearingsData
arrangedHearingsData <- selectedHearingsData %>%
  arrange(desc(id))
```

<br>

I find this function very useful when trying to inspect data. It provides as
outputs both the number of rows and columns, the names of the columns (variables)
in a dataset, and the structure of the each variable. All these could be 
achieved instead by running both `dim()`, `colnames()`, and `str()`. 

```{r}
# One thing that has not yet been covered in the course
glimpse(hearingsData)
```



***
Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


