---
title: "Alexander Genovese HW2 7.30.21"
description: |
  Alex Genovese's HW2 Revision 7.30.21
author:
  - name: Alex Genovese
date: 07-30-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
**Reading in the libraries**
```{r, include=FALSE}
library(tidyverse)
library(readxl)
```
I first read in the two libraries that I would need to read in the data and 
analyze it. Since the data I downloaded was an Excel doc, I used readxl 
library to do so.

**Reading in the data and cleaning it**
```{r}
GDPvsUnidata <- read_excel("../../_data/WorldBankGDPData.xlsx")
GDPvsUnidata[GDPvsUnidata == 0] <- NA
GDPdata <- GDPvsUnidata$`Millions of US Dollars`
Unidata <- GDPvsUnidata$`Number of Universities Worldwide in 2020 by Country`
```
I then read in the data into GDPvsUnidata. In order to have correct data into my 
analysis, I turned the 0 values into NA because it meant that the World Bank 
did not have data to support the GDP amount. I also created a vector for the 
GDP of each country based off Millions of US Dollars and another vector for the 
amount of Universities in each country.

**Looking at the data to see basic statistics**
```{r}
GDPdataMean <- mean(GDPdata, na.rm = TRUE)
GDPdataMedian <- median(GDPdata, na.rm = TRUE)
GDPdataRange <- max(GDPdata, na.rm = TRUE) - min(GDPdata, na.rm = TRUE)
UnidataMean <- mean(Unidata, na.rm = TRUE)
UnidataMedian <- median(Unidata, na.rm = TRUE)
UnidataRange <- max(Unidata, na.rm = TRUE) - min(GDPdata, na.rm = TRUE)
```
The data I wanted to analyze is does the amount of Universities impact GDP? I found the basic statistics of both the GDP data and the amount of Universities in each country. 

**Graphing the Data**
```{r}
ggplot(GDPvsUnidata, aes(x = Unidata, y = GDPdata)) + geom_point()
```
I graphed the data in a scatterplot to see how the amount of Universities impacts the GDP of a country. The data looks fairly linear, with the GDP rising with the amount of Universities in the country. There are two outliers in terms of the amount of Universities and how it impacts GDP, so I need to make sure that there is a causal relationship.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


