---
title: "Alexander Genovese HW2 7.27.21"
description: |
  Alex Genovese's HW2 Revision
author:
  - name: Alex Genovese
date: 07-27-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
**Reading in the libraries**
```{r, include=FALSE}
library(tidyverse)
library(readxl)
```
I first read in the two libraries that I would need to read in the data and 
analyze it. Since the data I downloaded was an Excel doc, I used readxl 
library to do so.

**Reading in the data and cleaning it**
```{r}
GDPdata <- read_excel("../../_data/WorldBankGDPData.xlsx")
GDPdata[GDPdata == 0] <- NA
GDPdataValue <- GDPdata$`Millions of US Dollars`
```
I then read in the data into GDPdata. In order to have correct data into my 
analysis, I turned the 0 values into NA because it meant that the World Bank 
did not have data to support the GDP amount. I also created a vector for the 
GDP of each country based off Millions of US Dollars.

**Looking at the data to see basic statistics**
```{r}
GDPdataMean <- mean(GDPdataValue, na.rm = TRUE)
print(GDPdataMean)
GDPdataMedian <- median(GDPdataValue, na.rm = TRUE)
print(GDPdataMedian)
GDPdataRange <- max(GDPdataValue, na.rm = TRUE) - min(GDPdataValue, na.rm = TRUE)
print(GDPdataRange)
```
The data I want to analyze is the large disparity in GDP from each country 
based off of whether they were first world, second world, or third world 
countries. The mean was pretty slanted, so I went with the median as a good 
idea of what the average is. I also found the range to show just how big the 
disparity in GDP is from the lowest GDP country to the highest. 

**Graphing the Data**
```{r}
class(GDPdata$Ranking)
class(GDPdataValue)
ggplot(GDPdata, aes(x = 'Ranking', y = GDPdataValue)) + geom_point()
```
I wanted to make sure that the data I was using was numerical to do a 
scatterplot. From there, I generated the scatterplot based off of ranking, 
with each number corresponding to the country (ie. 1 = US, 2 = China, 3 = 
Japan, etc.). With that graph, we can see just how slanted the data is towards 
the left (or lower GDP).
I am planning on separating these countries into how they're identified, 
first, second, or third world, by color and doing an analysis on how 
big the disparity is in 2021.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


