---
title: "Homework 4"
description: |
  Homework Four: Data Visualization
author:
  - name: Nathan Cote
date: 08-02-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---
```{r data setup, include=FALSE}
library(tidyverse)
library(readxl)
journalists <- read_excel("../../_data/journalists_Jul02.xlsx")
```
```{r recoded data, include=FALSE}
journalists_interaction <- journalists%>%
  group_by(INTERACTION) %>%
  filter(n() >=25)%>%
   mutate(Interplay = recode(INTERACTION, `15` = "Military vs Rioters", `16` = "Military vs Protesters", `17` = "Military vs Civilians", `27` = "Rebels vs Civilians", `37` = "Political Militia vs Civilians", `47` = "Communal Militia vs Civilians", `50` = "Sole Rioter Action", `55` = "Rioters vs Rioters", `56` = "Rioters vs Protesters", `57` = "Rioters vs Civilians", `58` = "Rioters vs Others", `60` = "Sole Protester Action", `68` = "Protesters vs Other", `78` = "Other Actor vs Civilians"))%>%
  ungroup()

journalists_joined <- (left_join(journalists, journalists_interaction))

event_count_per_interplay <- select(journalists_joined, EVENT_TYPE, Interplay)%>%
  group_by(Interplay)%>%
  count(EVENT_TYPE)%>%
  filter(n>=50)
event_count_per_interplay
```
# Introducing the Variables

The first visualization demonstrates the number of occurrences of interaction
types between various actors. It is essentially a count of my first variable.

```{r Interaction Count, echo=FALSE}
ggplot(data = na.omit(journalists_joined)) +
  geom_bar(mapping = aes(x = Interplay))+
  coord_flip()+
  labs(title = "Occurrences of Interaction Types Between Actors")
```
The second visualization demonstrates the number of occurrences of event types
among various actors. It is essentially a count of my second variable.

```{r Event Type Count, echo=FALSE}
ggplot(data = na.omit(journalists_joined)) +
  geom_bar(mapping = aes(x = EVENT_TYPE))+
  coord_flip()+
  labs(title = "Occurrences of Event Types Involving Actors")
```
# Putting Variables Together

The following visualizations demonstrate the occurrence of the variable 
Event Type among the variable, Interaction Type. First, I included every 
interaction type.

```{r Event Type and Interplay facet 1, echo=FALSE}
ggplot(data = na.omit(journalists_joined))+
  geom_bar(mapping = aes(x = EVENT_TYPE))+
  facet_wrap(~ Interplay)+
  coord_flip()+
  labs(title = "Occurrences of Event Types Among Interaction Types")
```

Because `Rioters vs Civilians` occurs so much more than the other interaction types,
it makes it difficult to visually interpret the lesser occurring interaction types.
I have decided to exclude `Rioters vs Civilians` in the visualization below to reduce that effect.

```{r Event Type and Interplay facet 2, echo=FALSE}
ggplot(data = na.omit(journalists_joined%>%filter(Interplay != "Rioters vs Civilians")))+
  geom_bar(mapping = aes(x = EVENT_TYPE))+
  facet_wrap(~ Interplay)+
  coord_flip()+
  labs(title = "Occurrences of Event Types Among Interaction Types",
       subtitle = "(Excludes Rioters vs Civilians interaction type)"
  )
```

The same difficulty to interpret the visualization is still present because `Military vs Rioters` has a similar, though to a lesser extent, effect as `Rioters vs Civilians`. I have decided to remove it in the visualization below.

```{r Event Type and Interplay facet 3, echo=FALSE}
ggplot(data = na.omit(journalists_joined%>%filter(Interplay != "Rioters vs Civilians" & Interplay != "Military vs Rioters")))+
  geom_bar(mapping = aes(x = EVENT_TYPE))+
  facet_wrap(~ Interplay)+
  coord_flip()+
  labs(title = "Occurrences of Event Types Among Interaction Types",
       subtitle = "(Excludes Rioters vs Civilians and Military vs Rioters interaction types)"
  )
```

# A Cleaner Visualization

After using facet wraps with my variables, I came accross an alternative approach
to data visualization which I find much cleaner and easier way to visually interpret. This alternative approach eliminates the "business" of the facet wrap visualization. 

```{r Event Type and Interplay fill 1, echo=FALSE}
ggplot(data = na.omit(journalists_joined))+
  geom_bar(mapping = aes(x = Interplay, fill = EVENT_TYPE),position = "dodge")+
  coord_flip()
```

```{r Event Type and Interplay fill 2, echo=FALSE}
ggplot(data = na.omit(journalists_joined%>%filter(Interplay != "Rioters vs Civilians")))+
  geom_bar(mapping = aes(x = Interplay, fill = EVENT_TYPE),position = "dodge")+
  coord_flip()+
  labs(title = "Occurrences of Event Types Among Interaction Types",
       subtitle = "(Excludes Rioters vs Civilians interaction type)"
  )
```

```{r Event Type and Interplay fill 3, echo= FALSE}
ggplot(data = na.omit(journalists_joined%>%filter(Interplay != "Rioters vs Civilians" & Interplay != "Military vs Rioters")))+
  geom_bar(mapping = aes(x = Interplay, fill = EVENT_TYPE),position = "dodge")+
  coord_flip()+
  labs(title = "Occurrences of Event Types Among Interaction Types",
       subtitle = "(Excludes Rioters vs Civilians and Military vs Rioters interaction types)"
  )
```
