---
title: "HW3"
description: |
 In this assignment I will be performing a univariate statistical analysis 
author:
  - name: Allie lyamuya
date: 07-28-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Import Data and Run Appropriate Librraries

```{r}
library(readr)
library(tidyverse)
library(tidyr)
library(ggplot2)

```


```{r}
# Import data
data <- read.csv("../../_data/digitalConnectivityData.csv")

# Description of the data
glimpse(data)

```

***


### Variable Description 

For this assignment, I'm interested in exploring two numeric variables from my
data, these are `internet_usage`, `gdp_percapita`. I will also use the variable
`population` for data transformation purposes, this is because both my variables
of interests have some relationship with the `population` variable. All three 
variables come from three different datasets obtained from the World Bank Open 
Data. Each variable include data from 2007 to 2017 (10 year period)

The variable `internet_usage` represent the percentage of people in a country
that are using the internet in a given year. The variable `gdp_percapita` 
represent the gross domestic product per population of a country in a 
given year. And the variable `population` represent the total population of a 
country in a given year. 


```{r}
# Describing the variable 
str(data$internet_usage)
str(data$population)
str(data$gdp_percapita)

summary(data$internet_usage)
summary(data$population)
summary(data$gdp_percapita)

```

***

### Data Cleaning

```{r}
# Creating a small dataset with my variable of interest
dataMini <- data %>%
  select(country_name, year, internet_usage, population, gdp_percapita)
head(dataMini)

# Check for missing values 
table(is.na(dataMini))

# Examine the variables with missing values
dataMini[rowSums(is.na(dataMini)) > 0, ]

```

<br>

There are **18** missing values in the small dataset `dataMini`. These missing 
values are found in the two variable, `gdp_percapita` and `internet_usage`. 
Because the dataset has a good number of observations (1947) and since I'm 
particularly interested to examine the relationship between the two variable 
with missing values, I will drop all observations with missing values to be able
to run statistical analyses more efficiently. 

```{r}
# Removing missing values 
dataMini1 <- na.omit(dataMini) 
table(is.na(dataMini1))

```

***

### Data Transformation

`internet_usage` is a percentage of a country population that was using the
the internet in a given year. I want to transform this percentage into an actual
portion of a country's population that was using the internet in a given year,
for the period between 2007 to 2017. To do that, I will use the variable 
`population` in computing `internet` (total internet usage in a country for a
given year)


```{r}
# Transforming variables
dataMini1 <- dataMini1 %>%
  rename("country" = "country_name", 
         "gdpc" = "gdp_percapita",
         "internet" = "internet_usage") %>%
  mutate(totalnet = (internet * population)/100) %>%
  arrange(country)

# Change the structure of `internet` to integer 
dataMini1$totalnet <- as.integer(dataMini1$totalnet)
```

### More on Data Tranformation

The World Bank categories countries' level of income into four groups, these are
low income, lower-middle income, upper-middle income, and high income. I will 
then categorized the `gdpc` variable into 4 categories based on the groups of 
percentiles in the dataset. Low income will be countries with `gdpc` less than 
the 25th percentile, between 25th and 50th percentile will be the low-middle 
income,between 50th and 75th will be the upper-middle income, and above 75th 
percentile will be the upper income. 

Moreover, for `internet`, I will group the level of internet usage across 
countries into three categories: Low, Moderate, and High. This grouping will be 
based on the groups of percentile and the median value of the variable `internet`.
Every level of usage below the median value will be grouped as Low, between 50th
and 75th percentile of use will be grouped as Moderate, and above the 75th perce
-ntile will be High. 

```{r}
# Calculate quantiles and summary of the gdpc variable
quantile(dataMini1$gdpc, na.rm = TRUE)

# Calculate quantiles and summary of the internet
summary(dataMini1$internet)
quantile(dataMini1$internet, na.rm = TRUE)

# Recode the data 
dataMini1 <- dataMini1 %>%
  mutate(gdpc_group = case_when(
    gdpc < 1673 ~ "Low Income",
    gdpc >= 1673 & gdpc <= 5586 ~ "Low-Middle Income",
    gdpc > 5586 & gdpc <= 20564 ~ "Upper-Middle Income",
    gdpc > 20564 ~ "Upper Income")) %>%
  mutate(internet_grp = case_when(
    internet < 37.3 ~ "Low",
    internet >= 37.3 & internet <= 67.09 ~ "Moderate",
    internet > 67.09 ~ "High"))
  
```

***

### Summary Description 

```{r}
str(dataMini1)
summary(dataMini1)

table(dataMini1$gdpc_group)
table(dataMini1$internet_grp)
```


***

### Visualization

```{r}
# Histograms 
ggplot(data = dataMini1) +
  geom_histogram(mapping = aes(x = internet), fill = "lightblue") +
  labs(title = "Histogram of Total Internet Usage") +
  theme_grey()

ggplot(data = dataMini1) +
  geom_histogram(mapping = aes(x = gdpc), fill = "lightpink") +
  ggtitle("Histogram of Countries' GDP per Capita")

```


***

### Exporting dataset

I export this subsetted data set so that I can be able to build on it in the 
next class assignment that focuses on visualization. After exporting the dataset
I renamed it and moved it to the right folder (_data). 

```{r}
#write.csv(dataMini1, 
         # "\\DACSS601SS22021\\_data\\dataMini1.csv", 
          #row.names = FALSE) 
```






***
 
Distill is a publication format for scientific and technical writing, native to 
the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


