---
title: "Homework One"
description: |
  A short description of the post.
author:
  - name: John Mannion

date: 07-28-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

# Now, let's check the first few rows. I'll let you remember how to do that!
myData <- read.csv("C:\\Users\\Thomas\\Desktop\\US-Executive-executive_orders_21.2.csv")

head(myData)

# How many rows of data are there in this dataset?

4459

# How many columns are there in the dataset?

18

# What are the column names?

#id eo_number president pres_party beg_term end_term signed_date year month day congress divided commemorative description pap_majortopic pap_subtopic majortopic subtopic

# Do one operation that requires creating a new object and doing something to that object

```{r, include=TRUE}
numb.exec.orders <- select(myData, 'president', 'id', 'pres_party')

ggplot(data = numb.exec.orders) +
  geom_point(mapping = aes(x = president, y = ))
```

# Do one thing to some portion of these data that hasn't been covered so far in the course materials. Be creative and have fun!

```{r, include=TRUE}
install.packages("devtools")
library(devtools)

devtools::install_github("ropensci/visdat")
library(visdat)

vis_dat(myData)
```
