---
title: Jimenez HW2
description: |
  Continuing with the MBTA ridership survey analysis
author:
- name: Sarah Jimenez
  url: http://umass.edu/sbs/dacss
date: 08-02-2021
output:
  distill::distill_article:
    self_contained: no
draft: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Picking up from HW1

I follow the same steps as in my previous post to import and start tidying my data. 

``` {r recap, message=FALSE}
all.rider <- read_csv("MBTA Systemwide Survey 2015-2017 Line Percentages.csv",
          na = c("", "NA"),
          col_types = cols(.default = "d", Mode = "c", Line = "c")
          )

rider <- select(all.rider,
  `Mode`,
  `Line`,
  `Household income: Less than $14,500`:`Count: Minority classification`,
  `Gender: Man`:`Count: Ability to understand English`)
  
rider <- rename(rider,
                mode = 'Mode',
                line = 'Line'
                )

rider <- rename(rider,
                hi.under14500 = 'Household income: Less than $14,500',
                hi.14500.28999 = 'Household income: $14,500 to $28,999',
                hi.29000.43499 = 'Household income: $29,000 to $43,499',
                hi.43500.57999 = 'Household income: $43,500 to $57,999',
                hi.58000.75999 = 'Household income: $58,000 to $75,999',
                hi.76000.108499 = 'Household income: $76,000 to $108,499',
                hi.108500.151999 = 'Household income: $108,500 to $151,999',
                hi.over1520000 = 'Household income: $152,000 or more',
                hi.pnts = 'Household income: Prefer not to say',
                hi.count = 'Count: Household income'
                )

rider <- rename(rider,
                       li.yes = 'Low-income classification: Yes',
                       li.no = 'Low-income classification: No',
                       li.count = 'Count: Low-income classification'
                       )

rider <- rename(rider,
                       race.AIAN = 'Race: American Indian or Alaska Native',
                       race.asian = 'Race: Asian',
                       race.black = 'Race: Black or African American',
                       race.NHPI = 'Race: Native Hawaiian or other Pacific Islander',
                       race.white = 'Race: White',
                       race.other = 'Race: Other',
                       race.pnts = 'Race: Prefer not to say',
                       race.count = 'Count: Race'
                       )

rider <- rename(rider,
                       lx.yes = 'Hispanic or Latino/Latina: Yes',
                       lx.no = 'Hispanic or Latino/Latina: No',
                       lx.pnts = 'Hispanic or Latino/Latina: Prefer not to say',
                       lx.count = 'Count: Hispanic or Latino/Latina'
                       )

rider <- rename(rider,
                       min.yes = 'Minority classification: Yes',
                       min.no = 'Minority classification: No',
                       min.count = 'Count: Minority classification'
                       )

rider <- rename(rider,
                       gen.m = 'Gender: Man',
                       gen.f = 'Gender: Woman',
                       gen.o = 'Gender: Other',
                       gen.pnts = 'Gender: Prefer not to say',
                       gen.count = 'Count: Gender'
                       )

rider <- rename(rider,
                       eng.always = 'Ability to understand English: Always',
                       eng.often = 'Ability to understand English: Often',
                       eng.some = 'Ability to understand English: Sometimes',
                       eng.never = 'Ability to understand English: Never',
                       eng.pnts = 'Ability to understand English: Prefer not to say',
                       eng.count = 'Count: Ability to understand English',
                       )

rider <- relocate(rider, line)

```

The result so far is this tibble:

```{r print}
print(rider)
```
There are 159 observations. The original dataset had 114 variables but I selected 41 of them and renamed them. 

## Further review and transformation

One research question for this data is: What is the proportion, by line and by mode, of riders that indicated on the survey that they are low-income, BIPO, women, and non-English fluent? 

Below I combine variables to create new variables that address the research question. (*Note there was no need to create a new variable for women as the **gen.f** variable already captures their responses). 

```{r mutate}

rider <- mutate(rider,
                hi.low = hi.under14500 + hi.14500.28999 + hi.29000.43499 + 
                hi.43500.57999,
                race.bipoc = race.AIAN + race.asian + race.black + race.NHPI + 
                race.other, 
                eng.nonfluent = eng.some + eng.never
                )

head(rider)
```

Now I can select for those new variables and trim the extra ones out. 

```{r}
rider <- select(rider,
  `line`,
  `mode`,
  `gen.f`,
  `hi.low`,
  `race.bipoc`,
  `eng.nonfluent`)

head(rider)
view(rider)

```

A number of lines do not have enough data to report and have missing values. 

