---
title: "HW4"
description: |
  Data Visualization for the variables of interest in the dataset I have chose
  for the final project. 
author:
  - name: Allie Lyamuya
date: 07-31-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Run the Libraries and Import the data

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
```


```{r}
# Load the data
dataSmall <- read.csv("../../_data/dataMini1.csv")
head(dataSmall)

# Explore the data
glimpse(dataSmall)
```

***
### Visualization: Tables and Crosstabs

```{r}
# Cross table for gdpc_group and internet_grp
dataSmall %>%
  select(gdpc_group, internet_grp) %>%
  table()


# Proportional table for gdpc_group and internet_grp
dataSmall %>%
  select(gdpc_group, internet_grp) %>%
  table() %>%
  prop.table() *100
```

* The cross table and proportional table above demostrate the distribution of 
different levels of countries' gdp per capita across different levels of 
internet usage by different countries across years. For example, in upper income 
countries, majority of the population are using the internet, that is there is
high internet usage across these countries population compared to a very small
percentage of low internet usage. 

* I chose this visualization approach because it is more straight forward and 
easy to grasp the relationship between the two variables in use. I will also 
explore other visualization approaches to get more nuances of the relationship 
between the internet usage and countries' income level. 


***
### Visualization: GGPlot - Univariate and Bivariate 

```{r}
# Univarariate: Histograms
ggplot(data = dataSmall) +
  geom_histogram(mapping = aes(x = internet), fill = "chocolate1") +
  labs(title = "Histogram of the Percentage of Internet Usage in Countries") +
  theme_grey()

ggplot(data = dataSmall) +
  geom_histogram(mapping = aes(x = gdpc), fill = "aquamarine4") +
  ggtitle("Histogram of Countries' GDP per Capita")

# Univariate: Bar graphs
ggplot(data = dataSmall) +
  geom_bar(mapping = aes(x = internet_grp), fill = "darkgoldenrod1") +
  labs(title = "Bar Graph of the Level of Internet Usage in Countries")

ggplot(data = dataSmall) + 
  geom_bar(mapping = aes(x = gdpc_group), fill = "blueviolet") +
  labs(title = "Bar Graph of the Levels of Countries' GDP per Capita")

```

i. The above two histograms represent the distribution of the numerical 
variables, `internet` and `gdpc`. Based on the histogram of `internet`, lower
percentage of internet usage has a large count compared to high percentage of
internet usage. Besides, the `gdpc` histogram reveals a high number of countries
with low level of income compared to numbers of high income countries. 

ii. The two bar graphs represent the distribution of the categorical variables
`internet_grp` and `gdpc_group`. Based on the `internet_grp` bar graph, majority
of countries seem to have low level of internet usage across their population.
Moreover, the `gdpc_grp` bar graph reveals that the distribution of income level
by different groups of countries seem to have the same proportion in the dataset.
This can be attributed to the creation of the variable `gdpc_grp` based on the
four levels of the percentiles (0 - 25%, 25% - 50%, 50% - 75%, and 75% - 100%). 

iii. I chose histograms and bar graphs because they are good visualization
method for visualizing the distribution of univariate numeric variable and 
categorical variables across time. 

***

```{r}
# Bivariate: Scatterplot for `internet` and `gdpc`
ggplot(data = dataSmall) +
  geom_point(mapping = aes(x = gdpc, y = internet, color = population)) +
  labs(title = "Scatterplot of GDP per Capita and Internet Usage by Countries")
  
# Bivariate: Boxplot for internet and groups of gdpc
ggplot(data = dataSmall, mapping = aes(x = gdpc_group, y = internet)) +
  geom_boxplot(color = "mediumorchid4") +
  labs(title = "Boxplot for Levels of GDP per Capita and Internet Usage")

# Bivariate: Bar Graph for levels of internet usage and levels of gdp/capita 
ggplot(data = dataSmall) +
  geom_bar(mapping = aes(x = gdpc_group, fill = internet_grp), 
           position = "dodge") +
  labs(title = "Bar Graph of GDP per Capita by Levels of Internet Usage")

```


i. The scatterplot above shows the positive relationship between countries'
gdp per capita (numeric variable) and the percentage of internet usage (numeric
variable). That is, as `gdpc` increase there seem to be an increase in 
countries' internet usage. I chose scatterplot because this form of 
visualization is best to represent the relationship between two numeric 
variables. 

- I tried to use `geom_bar` here and it didn't seem to work with two continuous 
variables. 


ii. The boxplot above shows the relationship between different levels of
countries' gdp per capita (categorical variable) and the percentage of internet
usage by countries (numeric variable). In the Upper Income group, the boxplot
display the presence of many outliers--portion of countries' populations with
lower percentage of internet usage. Also, the Low Income group have some 
outliers--portion of countries' population with a high internet usage. Besides, 
the distribution of the maximum values, 1st and 3rd quartiles across the four 
groups reveals the positive relationship between the internet usage and levels 
of income different countries belong to. 

- I choose boxplot because it is great at visualizing the distribution of data
when examining the relationship between a numeric variable and categorical
variable. I could also use `geom_violin` but I decided not to use it here. 


iii. The bar graph above displays the distribution of levels of gdp per capita 
of varied groups of countries (categorical variable) while also showing the 
distribution of internet usage based on varied levels across countries 
(categorical variable). In each group, from Low to Upper Income, there are 3 
groups of internet usage from Low to Moderate to High. 

- I chose this visualization form because it allows for the representation of 
distribution of categorical variable while also allow the option to embed 
another categorical variable in the same graph to display the distribution of
both variables across time. I could have used line graph here but that would 
likely limit the ability to display another variable at the same time. 


iv. **Limitation**

The correlated relationship between variables that both scatterplot, boxplot, 
and bar graph have revealed above (based on the visualizations) don't 
necessarily say anything about causality. That is, while there seem to be a 
revealed relationship between the internet usage and the countries levels of 
income, it is not that clear from the visualization above how these two 
variables affect one another. I would like to perform other, perhaps more 
advanced visualization techniques that can explain how the two variable of 
interest from my dataset affect one another. 


***

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


